Assessing Music Trends of Beyoncé and Coldplay

### 2024-11-14 paper/paper.qmd
update script

---
title: "Assessing Music Trends of Beyoncé and Coldplay"
subtitle: "Song Duration and Valence Analysis Using Spotify Data"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/spotify)."
date: today
date-format: long
abstract: "This paper analyzes the audio features of songs from two globally renowned artists, Beyoncé and Coldplay, using Spotify's API. By focusing on song duration and valence (musical positivity), the study compares trends in their music over time. The results show that Beyoncé's songs vary more in length and tend to have higher valence, while Coldplay's tracks are more consistent in duration but exhibit a more melancholic tone."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(spotifyr)
library(lubridate)
library(ggplot2)

coldplay <- read_rds(here::here("data/01-raw_data/coldplay.rds"))
beyonce <- read_rds(here::here("data/01-raw_data/beyonce.rds"))
```

@@

```{r}
#| label: fig-duration-beyonce
#| fig-cap: Duration of Songs of Beyonce
#| echo: false
#| warning: false
#| message: false

beyonce <- as_tibble(beyonce)

beyonce |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

```{r}
#| label: fig-duration-coldplay
#| fig-cap: Duration of Songs of Coldplay
#| echo: false
#| warning: false
#| message: false

coldplay <- as_tibble(coldplay)

coldplay |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

@@

rbind(beyonce, coldplay) |>
  select(artist_name, album_release_date, valence) |>
  mutate(album_release_date = ymd(album_release_date)) |> 
  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

>>>>

---
title: ""
subtitle: ""
author: 
  - Andy Jiang
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/grocery-price-analysis.git)."
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(DBI)
library(RSQLite)
library(ggplot2)

# Connect to the SQLite database
con <- dbConnect(RSQLite::SQLite(), "/Users/y4nxunj/Downloads/hammer-2-processed.sqlite")  # Update path

# Run the join query from '01_join_raw_and_product.sql' to get combined data
query <- "
  SELECT 
      r.nowtime, 
      p.vendor, 
      r.product_id, 
      p.product_name, 
      p.brand, 
      r.current_price, 
      r.old_price, 
      p.units, 
      r.price_per_unit, 
      r.other 
  FROM 
      raw AS r
  INNER JOIN 
      product AS p ON r.product_id = p.id
  ORDER BY 
      r.nowtime;
"
data <- dbGetQuery(con, query)

# Disconnect from the database
dbDisconnect(con)

# Now you have 'data' in R and can use it for visualizations, e.g., average price by vendor
avg_price_by_vendor <- aggregate(current_price ~ vendor, data, mean)

ggplot(avg_price_by_vendor, aes(x = vendor, y = current_price, fill = vendor)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Price by Vendor",
       x = "Vendor", y = "Average Price") +
  theme_minimal() +
  theme(legend.position = "none")

```

@@

```{r}
#| label: fig-duration-coldplay
#| fig-cap: Duration of Songs of Coldplay
#| echo: false
#| warning: false
#| message: false
# Filter for rows where products were on sale
sales_data <- subset(data, !is.na(old_price))

# Count sales occurrences by vendor
sales_count <- as.data.frame(table(sales_data$vendor))
colnames(sales_count) <- c("vendor", "sale_count")

# Plot frequency of sales by vendor
ggplot(sales_count, aes(x = vendor, y = sale_count, fill = vendor)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Sales by Vendor",
       x = "Vendor", y = "Sale Count") +
  theme_minimal() +
  theme(legend.position = "none")

```

@@

# Improved box plot with fewer jitter points and adjusted transparency and size
set.seed(42)  # For reproducible sampling
sampled_data <- data[sample(nrow(data), size = min(5000, nrow(data))), ]  # Sample 5000 points or fewer

ggplot(data, aes(x = vendor, y = current_price, fill = vendor)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Hide outliers in box plot
  geom_jitter(data = sampled_data, width = 0.2, alpha = 0.1, size = 0.5, color = "black") +  # Sampled points with low transparency
  scale_y_continuous(trans = 'log10', limits = c(1, 100)) +  # Log scale and limit y-axis
  labs(title = "Price Distribution Across Vendors",
       x = "Vendor", y = "Current Price (Log Scale)") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
  scale_fill_brewer(palette = "Set3")  # Use a color palette for better presentation

```

```{r}
# Plot distribution of prices by vendor
# Calculate average price per vendor
avg_price_by_vendor <- aggregate(current_price ~ vendor, data, mean, na.rm = TRUE)

# Plot average price by vendor
ggplot(avg_price_by_vendor, aes(x = vendor, y = current_price, fill = vendor)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Price by Vendor",
       x = "Vendor", y = "Average Price") +
  theme_minimal() +
  theme(legend.position = "none")

```
